---
title: "Homework 4"
author: "Carly Greutert"
date: "5 March 2023"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(knitr)
eval=TRUE
knitr::opts_chunk$set(
  eval=eval,
  echo=TRUE, 
  cache=FALSE, 
  fig.width=5, 
  fig.height=5,
  fig.align='center')
r = function(x, digits=2){ round(x, digits=digits) }
indent1 = '    '      
indent2 = paste(rep(indent1, 2), collapse='')
library(tidyverse)
library(reshape2)
library(magrittr)
library(testthat)
library(cmdstanr)
```

**Note:** If you are working with a partner, please submit only one
homework per group with both names and whether you are taking the course
for graduate credit or not. Submit your Rmarkdown (.Rmd) and the
compiled pdf on Gauchospace.

### Problem 1. Frequentist Coverage of The Bayesian Posterior Interval.

In the "random facts calibration game" we explored the importance and difficulty of well-calibrated
prior distributions by examining the calibration of subjective
intervals. Suppose that $y_1,..,y_n$ is an IID sample from a
$Normal(\mu, 1)$. We wish to estimate $\mu$.

**1a.** For Bayesian inference, we will assume the prior distribution
$\mu \sim Normal(0,\frac{1}{\kappa_0})$ for all parts below. Remember,
from lecture that we can interpret $\kappa_0$ as the pseudo-number of
prior observations with sample mean $\mu_0 = 0$. State the posterior
distribution of $\mu$ given $y_1,..,y_n$. Report the lower and upper
bounds of the $95\%$ quantile-based posterior credible interval for
$\mu$, using the fact that for a normal distribution with standard
eviation $\sigma$, approximately $95\%$ of the mass is between
$\pm 1.96\sigma$. \newline

$\mu | Y$ ~ N($\mu_n$, $\tau^2_n$) where 
$\mu_n = \frac{\frac{1}{\frac{1}{\kappa_0}} * 0 + \frac{n}{1} * \bar{y}}{\frac{1}{\frac{1}{\kappa_0}} + \frac{n}{1}}$ = $\frac{\kappa_0  * 0 + n\bar{y}}{\kappa_0 + n}$ = $\frac{n\bar{y}}{\kappa_0 + n}$ \
and $\tau^2_n = \frac{1}{\frac{1}{\frac{1}{\kappa_0}} + \frac{n}{1}}$ = $\frac{1}{\kappa_0 + n}$ \
so $\mu | Y ~ N(\frac{n\bar{y}}{\kappa_0 + n}, \frac{1}{\kappa_0 + n})$ \
The lower bound of the $95\%$ quantile-based posterior credible interval is $\frac{n\bar{y}}{\kappa_0 + n} - 1.96 * \sqrt{\frac{1}{\kappa_0 + n}}$ \
The upper bound is $\frac{n\bar{y}}{\kappa_0 + n} + 1.96 * \sqrt{\frac{1}{\kappa_0 + n}}$

**1b**. Plot the length of the posterior credible interval as a function
of $\kappa_0$, for $\kappa_0 = 1, 2, ..., 25$ assuming $n=10$. Report
how this prior parameter effects the length of the posterior interval
and why this makes intuitive sense.

```{r}
# Use 'interval_length' to store lengths of credible intervals
interval_length <- numeric()
for (kappa in 1:25){
  interval_length[kappa] = 2 * 1.96 * sqrt(1/(kappa + 10))
}

ggplot(tibble(interval_length), aes(x= 1:25, y = interval_length)) + 
  xlab("Value of kappa_0") + ylab("Length of Interval") +
  geom_line()

```

```{r}
. = ottr::check("tests/q1b.R")
```
As the prior parameter, $\kappa_0$ increases, the posterior credible interval decreases. This makes intuitive sense because having a higher value of the prior parameter typically signals confidence/a better prediction about $\mu$ and therefore we can shorten the interval to a more accurate guess. 

**1c**. Now we will evaluate the *frequentist coverage* of the posterior
credible interval on simulated data. Generate 1000 data sets where the
true value of $\mu=0$ and $n=10$. For each dataset, compute the
posterior $95\%$ interval endpoints (from the previous part) and see if
it the interval covers the true value of $\mu = 0$. Compute the
frequentist coverage as the fraction of these 1000 posterior 95%
credible intervals that contain $\mu=0$. Do this for each value of
$\kappa_0 = 1, 2, ..., 25$. Plot the coverage as a function of
$\kappa_0$.  Store these 25 coverage values in vector called `coverage`.

```{r}

## Fill in the vector called "coverage", which stores the fraction of intervals containing \mu = 0 for each kappa.
covers <- c()
coverage <- numeric(25)
for (kappa in 1:25){
for (i in 1:1000){
  yi = rnorm(n = 10, mean = 0, sd = 1)
  lower = (10*mean(yi))/(kappa + 10) - 1.96 * sqrt(1/(kappa + 10))
  upper = (10*mean(yi))/(kappa + 10) + 1.96 * sqrt(1/(kappa + 10))
  covers[i] = between(0, lower, upper)
}
  coverage[kappa] = sum(covers)/1000
  }
ggplot(tibble(coverage), aes(x= 1:25, y = coverage)) + 
  xlab("Value of kappa_0") + ylab("Fraction of Intervals Containing Mu = 0") +
  geom_line()
```

```{r}
. = ottr::check("tests/q1c.R")
```

**1d.** Repeat 1c but now generate data assuming the true $\mu=1$. Again, store these 25 coverage values in vector called `coverage`.

```{r}

## Fill in the vector called "coverage", which stores the fraction of intervals containing \mu = 1 for each kappa.
coverage <- numeric(25)
covers <- c()

for (kappa in 1:25){
for (i in 1:1000){
  yi = rnorm(n = 10, mean = 1, sd = 1)
  lower = (10*mean(yi))/(kappa + 10) - 1.96 * sqrt(1/(kappa + 10))
  upper = (10*mean(yi))/(kappa + 10) + 1.96 * sqrt(1/(kappa + 10))
  covers[i] = between(1, lower, upper)
}
  coverage[kappa] = sum(covers)/1000
  }
ggplot(tibble(coverage), aes(x= 1:25, y = coverage)) + 
  xlab("Value of kappa_0") + ylab("Fraction of Intervals Containing Mu = 0") +
  geom_line()
```

```{r}
. = ottr::check("tests/q1d.R")
```

**1e**. Explain the differences between the coverage plots when the true
$\mu$ = 0 and the true $\mu = 1$. For what values of $\kappa_0$ do you
see closer to nominal coverage (i.e. 95%)? For what values does your
posterior interval tend to overcover (the interval covers the true value
more than 95% of the time)? Undercover (the interval covers the true
value less than 95% of the time)? Why does this make sense?

When $\mu$ = 0, the higher value of $\kappa_0$ leads to a higher coverage value whereas when $\mu$ = 1, the reverse is true, the higher value of $\kappa_0$ leads to a lower rate of coverage. When $\mu$ = 0, all values of $\kappa_0$ lead to above 95% coverage, so they all overcover. When $\mu$ = 1, $\kappa_0$ = 1, 2 leads to almost 95% coverage, but the rest undercover. This makes sense because the variance is equivalent to $\frac{1}{\kappa_0}$, so the variance shrinks towards zero as we get more confident in our prior, so our interval is less likely to cover 1 as much as it will 0. 


### Problem 2. Goal Scoring in the Women's World Cup

Let's take another look at scoring in soccer.  The Chinese Women's soccer team recently won the AFC Women's Asian Cup. Suppose you are interested in studying the World Cup performance of this soccer team. Let $\lambda$ be the be the average number of goals scored by the team. We will analyze $\lambda$ using the Gamma-Poisson model where data $Y_i$ is the observed number of goals scored in the $i$th World Cup game, ie. we have $Y_i|\lambda \sim Pois(\lambda)$.  _A priori_, we expect the rate of goal scoring to be $\lambda \sim Gamma(a, b)$. According to a sports analyst, they believe that $\lambda$ follows a Gamma distribution with $a = 1$ and $b = 0.25$. 

**2a.** 
Compute the theoretical posterior parameters a, b, and also the posterior mean. 

```{r}
y <- c(4, 7, 3, 2, 3) # Number of goals in each game

post_a <- sum(y) + 1
post_b <- length(y) + 0.25
post_mu <- post_a/post_b
```

```{r}
. = ottr::check("tests/q2a.R")
```

**2b.** 
Create a new Stan file by selecting "Stan file" and name it `women_cup.stan`. Encode the Poisson-Gamma model in Stan.  Use `cmdstanr` to report and
estimate the posterior mean of the scoring rate by computing the sample
average of all Monte Carlo samples of $\lambda$.

```{r}

## Create "women_cup.stan" yourself and fill in the model
soccer_model <- cmdstan_model("women_cup.stan")

## This fits the model to data y 
## All parameter samples are stored in a data frame called "samples"
stan_fit <- soccer_model$sample(data=list(N = length(y), Y = y), refresh=0, show_messages = FALSE)
samples <- stan_fit$draws(format="df")

## Compute the posterior mean of the lambda samples

post_mean <- stan_fit$summary()$mean[2]

```

```{r}
. = ottr::check("tests/q2b.R")
```
**2c.**
Create a histogram of the Monte Carlo samples of $\lambda$ and add a line showing the theoretical posterior of density of $\lambda$. Do the Monte Carlo samples coincide with the theoretical density?

```{r}
f1 <- function(x){dgamma(x, post_a, post_b)}
ggplot(tibble(samples$lambda), aes(x = samples$lambda)) + geom_histogram(aes(y = ..density..)) +  stat_function(fun = f1, color = 'blue')
```

**2d.**
Use the Monte Carlo samples from Stan to compute the mean of predicative posterior distribution to estimate the distribution of expected goals scored for next game played by the Chinese women's soccer team.

```{r}
pred_mean <- mean(rpois(n = length(samples$lambda), lambda = samples$lambda))
```

```{r}
. = ottr::check("tests/q2d.R")
```

### Problem 3. Bayesian inference for the normal distribution in Stan.


Create a new Stan file and name it `IQ_model.stan`. We will make some basic modifications to the
template example in the default Stan file for this problem. Consider the
IQ example used from class. Scoring on IQ tests is designed to yield a
N(100, 15) distribution for the general population. We observe IQ scores
for a sample of $n$ individuals from a particular town,
$y_1, \ldots y_n \sim N(\mu, \sigma^2)$. Our goal is to estimate the
population mean in the town. Assume the
$p(\mu, \sigma) = p(\mu \mid \sigma)p(\sigma)$, where
$p(\mu \mid \sigma)$ is $N(\mu_0, \sigma/\sqrt{\kappa_0})$ and
$p(\sigma)$ is Gamma(a, b). Before you administer the IQ test you
believe the town is no different than the rest of the population, so you
assume a prior mean for $\mu$ of $\mu_0 = 100$, but you aren't to sure
about this a priori and so you set $\kappa_0 = 1$ (the effective number
of pseudo-observations). Similarly, a priori you assume $\sigma$ has a
mean of 15 (to match the intended standard deviation of the IQ test) and
so you decide on setting $a=15$ and $b=1$ (remember, the mean of a Gamma
is a/b). Assume the following IQ scores are observed:

```{r}
y <- c(70, 85, 111, 111, 115, 120, 123)
n <- length(y)
```

**3a**. Make a scatter plot of the posterior distribution of the mean,
$\mu$, and the precision, $1/\sigma^2$. Put $\mu$ on the x-axis and
$1/\sigma^2$ on the y-axis. What is the posterior relationship between
$\mu$ and $1/\sigma^2$? Why does this make sense? *Hint:* review the
lecture notes. 

```{r}
normal_stan_model <- cmdstan_model("IQ_model.stan")

# Run rstan and extract the samples
stan_fit <- normal_stan_model$sample(data=list(N = n, y = y, mu0 = 100, k0 = 1), refresh=0, show_messages = FALSE)
samples <- stan_fit$draws(format="df")

mu_samples <- samples$mu
sigma_samples <- samples$sigma
precision_samples <- 1/((sigma_samples)^2)

## Make the plot
tibble(Mean = mu_samples, Precision=precision_samples) %>% 
 ggplot() +
 geom_point(aes(x=Mean, y=Precision)) +
 theme_bw(base_size=16)
```

```{r}
. = ottr::check("tests/q3a.R")
```
I notice that when precision is high, there is a large sampling variance and when it is low, there is low sampling variance. This makes intuitive sense because precision is the reciprocal of the variance, so when variance is high, the precision will be low, they have an inverted relationship. To continue, a lower sampling variance is indicative of higher confidence in $\mu$ so their values are spread further apart. 


**3b**. You are interested in whether the mean IQ in the town is greater
than the mean IQ in the overall population. Use Stan to find the
posterior probability that $\mu$ is greater than 100.

```{r}
mean(mu_samples > 100)
```

Out of all the generated posterior $\mu$ samples, it appears that ~76.9% of the town IQ is greater than the mean of the overall population. 

**3c.** The [coefficient of variation](https://en.wikipedia.org/wiki/Coefficient_of_variation), $c_v=\sigma/\mu$ is defined as the standard deviation over the mean. Make a histogram of $p(c_v\mid y)$ from Monte Carlo samples and report the posterior mean and the lower and upper endpoints of the 95\% quantile based interval.

```{r}
cv <- c()
for (s in 1:1000){
  mu <- rgamma(n, 15 + sum(y), 1 + n)
  tilde_Y <- rpois(n, mu)
  cv <- append(cv, sd(tilde_Y)/mean(tilde_Y))
}
ggplot(tibble(cv), aes(x = cv)) + geom_histogram(aes(y = ..density..))
mean(cv)
quantile(cv, 0.05)
quantile(cv, 0.95)
```

